### 2016-11-21 local-capability-machine.tex
Some remarks and questions...

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\begin{description}
\item[\texttt{call f(r)}] Moves the argument to a register (if there are more arguments, then they are moved to the appropriate registers). Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return). Erases all registers but the argument register(s), $\pcreg$, the $\malloc$ register (if there is one), and the return register. Jumps to $\texttt{f}$.
\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}

\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop r}] Pops the top word of the stack by loading it to regiser $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall f(r)}] This call assumes the presence of a stack! Push the code that restores the stackpointer, $\pcreg$ and $r_0$. Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer. Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$. Restrict the stackpointer to the ``unused part''. Push the arguments to the stack and jump to \texttt{f}.
  \begin{description}
  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code.
  \end{description}
Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
\end{description}

@@

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
ML-like program we had before:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack
\begin{verbatim}
f1: malloc l
    hstore l 1
    call adv()
    assert l 1
    halt
1f:
\end{verbatim}
\begin{lemma}[Correctness lemma for \texttt{f1}]
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]}
\end{align*}
and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
then
\[
  h'(\flag) = 0
\]  
\end{lemma}

>>>>

\section{Macros}
In order to write readable example programs, we provide macros that can be programmed using the instruction set given in the formalisation.
\dominique{how is the return capability constructed: not on the stack, I presume, so this means mallocing a block of memory for the activation record?}
\dominique{Similarly: how is malloc invoked?  Do we trust malloc enough to not encapsulate ourselves from it, i.e. provide an rx return capability and use callee-save registers?}
\dominique{what does ``fetch the capability ...'' mean?}
\dominique{Wouldn't it be more clear to provide call with two explicit lists of registers: those which need to be stored, and those which are provided as arguments (i.e. which do not need to be erased)?}
\dominique{Perhaps you could also provide an explicit syntax for ``undefined symbols'' that should be filled in by a linker?}
\begin{description}
\item[\texttt{call f(r)}] Moves the argument to a register (if there are more arguments, then they are moved to the appropriate registers). Sets up a return capability in $r_0$ (this includes storing existing capabilities, so they can be read upon return). Erases all registers but the argument register(s), $\pcreg$, the $\malloc$ register (if there is one), and the return register. Jumps to $\texttt{f}$.
\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}

\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop r}] Pops the top word of the stack by loading it to regiser $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall f(r)}] This call assumes the presence of a stack! Push the code that restores the stackpointer, $\pcreg$ and $r_0$. Push the code return point, $r_0$ (the continuation for this execution), and the stack pointer \dominique{and other registers?}. Set up a continuation for \texttt{f} that is make the stack pointer point to the ``restore code'' and restrict it to an enter capability and move it to $r_0$. Restrict the stackpointer to the ``unused part''. Push the arguments to the stack and jump to \texttt{f}.
  \begin{description}
  \item[Restore code] Restores the stack capability, the continuation zeros the part of the stack passed to \texttt{f} and finally jumps to the return point in code. \dominique{restores other registers too, right?}
  \end{description}
Figure \ref{fig:stack-before-call}-\ref{fig:stack-after-call} illustrates what the stack would look like at different stages of the scall.
\end{description}

@@

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
ML-like program we had before:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack
\begin{verbatim}
f1: malloc l
    hstore l 1
    call adv()
    assert l 1
    halt
1f:
\end{verbatim}
\begin{lemma}[Correctness lemma for \texttt{f1}]
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
  \end{align*}
if \dominique{$\hs_f$ \emph{and} $\hs_{f1}$?}
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus h_{\malloc} \uplus \hs_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
\end{align*}
and $hs_\flag(\flag) = 0$. $hs_g$ contains the code of $g$ \dominique{do you mean f1?} including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
then
\[
  h'(\flag) = 0
\]  
\end{lemma}
